---
title: "Creating a Wrapper for New Functionality and Registering It"
author: Dario Strbenac <br>
        The University of Sydney, Australia.
output: 
    BiocStyle::html_document:
        css: style.css
        toc: false
---

## Introduction

There might be a new transformation / selection / modelling algorithm that's not a part of ClassifyR but might be useful to others. This guide explains the necessary steps to perform before making a pull request to get new functionality added.

The core framework dispatches the data as a `DataFrame` object, so the new function should be able to accept such a variable type as its first argument.

## Steps to Add a New Function

1. Define an ordinary R function and attach a name to it. This will be useful in automated annotation of modelling results later. A basic outline is:

```{r}
newFancyClassifier <- function(measurementsTrain, classesTrain, alpha, ..., verbose = 3)
{
  # Build a model.  
}
attr(newFancyClassifier, "name") <- "newFancyClassifier"
```

Each interface follows the contract in R/interfaceTemplate.R: the arguments it receives and the format of the predictions it returns.

2. Add the new file to the `Collate` field of the DESCRIPTION file.

3. Register the function with one entry in R/registry.R, inside `.buildRegistry`. The entry gives the keyword that users type, the training and prediction functions, the name shown in results and plots, a one-sentence description for `available()`, the kinds of outcome the method accepts and, if the classifier has tuning parameters, a small range of values to try when `tuneParams` is `"auto"`. For example,

```{r, eval = FALSE}
.registerClassifier("fancy", newFancyClassifier, newFancyPredictor, displayName = "Fancy Classifier",
                    description = "Very good classifier that predicts amazingly well.",
                    outcomes = "classes", tunePresets = list(alpha = c(0.1, 0.5, 1)))
```

A feature selection method is registered with `.registerSelection`, giving its ranking function instead.

4. Run the package's tests. A test fits and predicts every registered keyword on small data sets and checks the format of the predictions.

5. Open the vignette in vignettes/ClassifyR.Rmd and look for the section heading #### Provided Methods for Feature Selection and Classification. Add the new function to the appropriate table.

Now that the new function is defined and described, users will be able to discover it using `available()` at the R command line and results will automatically store nice names for the functions used and they will automatically be able to be shown in plots.